---
title: "App-Package Structure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{App-Package Structure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

## What is a Shiny app-package?

A **Shiny app-package** is an R package that also contains a Shiny application.
It satisfies two requirements simultaneously:

- It has a valid `DESCRIPTION` file, so R treats it as a package.
- It contains a `launch_app()` function (and an `app.R` entry point) that
  runs the app.

This combination gives a Shiny app everything a package provides: proper
dependency management, a documentation system, unit tests, and vignettes —
without any extra tooling.

## File layout

```
movexplR6/
├── DESCRIPTION              # package identity and dependencies
├── NAMESPACE                # exported symbols (roxygen2-generated)
├── movexplR6.Rproj          # RStudio project file
├── app.R                    # thin entry point for runApp() / Run App button
├── R/
│   ├── MovieData.R          # R6 class (database + filtering)
│   ├── utils.R              # axis_vars constant
│   ├── mod_filters.R        # filter inputs module
│   ├── mod_plot.R           # scatter plot module
│   ├── movies_ui.R          # top-level UI function
│   ├── movies_server.R      # top-level server function
│   └── launch_app.R         # launch_app() entry point
├── inst/
│   └── extdata/
│       └── movies.db        # bundled SQLite database
├── tests/
│   ├── testthat.R           # test runner
│   └── testthat/
│       ├── helper.R         # shared setup (db_path)
│       ├── test-utils.R
│       ├── test-MovieData.R
│       ├── test-mod_filters.R
│       └── test-mod_plot.R
└── vignettes/
    ├── movexplR6.Rmd        # introduction and usage
    ├── r6-class.Rmd         # R6 class design
    ├── shiny-modules.Rmd    # module architecture
    └── app-package-structure.Rmd  # this vignette
```

## `DESCRIPTION`

`DESCRIPTION` is what makes a directory an R package. For a Shiny app-package
the most important fields are `Imports` and `Suggests`.

```{r description, eval = FALSE}
Package: movexplR6
Version: 0.0.0.9000
Type: Package
Title: Movie Explorer Shiny App-Package
Imports:
    R6,
    shiny,
    bslib,
    dplyr,
    ggplot2,
    plotly,
    DBI,
    RSQLite,
    dbplyr,
    logger
Suggests:
    testthat (>= 3.0.0),
    knitr,
    rmarkdown
Config/testthat/edition: 3
VignetteBuilder: knitr
```

**`Imports`** lists packages that must be installed for the app to run.
`devtools::install()` installs them automatically. Inside package code, these
packages are called with their namespace prefix (`shiny::reactive()`,
`dplyr::filter()`) rather than `library()`.

**`Suggests`** lists packages only needed for testing or building documentation.
They are not installed when someone `install.packages()` the package.

## `NAMESPACE`

`NAMESPACE` lists the symbols the package makes available to users. It is
generated by `devtools::document()` from `@export` tags in the roxygen2
comments, so you should never edit it by hand.

```{r namespace, eval = FALSE}
# Generated by roxygen2: do not edit by hand

export(MovieData)
export(axis_vars)
export(launch_app)
export(mod_filters_server)
export(mod_filters_ui)
export(mod_plot_server)
export(mod_plot_ui)
export(movies_server)
export(movies_ui)
```

Exporting module functions and `MovieData` means users can call them directly —
useful for composing the package with other packages or for testing.

## The `R/` directory

Each file in `R/` contains one logical unit. Keeping files single-purpose makes
them easier to locate and test.

| File | Contents |
|---|---|
| `MovieData.R` | R6 class; database connection, data loading, filtering |
| `utils.R` | `axis_vars` named vector shared by modules |
| `mod_filters.R` | `mod_filters_ui()` + `mod_filters_server()` |
| `mod_plot.R` | `mod_plot_ui()` + `mod_plot_server()` |
| `movies_ui.R` | `movies_ui()` — top-level layout |
| `movies_server.R` | `movies_server()` — wires modules to MovieData |
| `launch_app.R` | `launch_app()` — calls `shiny::shinyApp()` |

No `global.R`, no `ui.R`, no `server.R`. Everything that was in those files
in the original `051-movie-explorer` example now lives in named, documented,
testable functions inside `R/`.

## `inst/extdata/`

Files under `inst/` are copied verbatim into the installed package. `extdata/`
is the conventional subdirectory for external data files (the `Writing R
Extensions` manual recommends it for non-R data).

`system.file()` resolves the path at runtime whether the package is installed
or loaded with `devtools::load_all()`:

```{r system-file, eval = FALSE}
db_path <- system.file("extdata/movies.db", package = "movexplR6")
```

This means `MovieData$new()` never needs a hard-coded file path.

## `app.R` — the entry point

```{r app-r, eval = FALSE}
pkgload::load_all()
launch_app()
```

`app.R` does two things:

1. `pkgload::load_all()` simulates `devtools::install()` in the current
   session, making all exported functions available without a formal install step.
2. `launch_app()` calls `shiny::shinyApp(ui = movies_ui(), server = movies_server)`.

This two-liner is the minimum needed for the RStudio **Run App** button and for
`shiny::runApp(".")` to work. All real logic lives in `launch_app()`, not here.

## `launch_app()`

```{r launch-app, eval = FALSE}
launch_app <- function(...) {
  shiny::shinyApp(
    ui     = movies_ui(),
    server = movies_server,
    ...
  )
}
```

`launch_app()` is the package's public API for running the app. The `...`
argument passes additional options to `shiny::shinyApp()` (e.g., `options =
list(port = 4321)`).

Note that `ui` receives the **result** of `movies_ui()` (evaluated once), while
`server` receives the **function object** `movies_server` (not called here).
Shiny calls the server function once per session; the UI is built once at startup.

## Tests

The `tests/testthat/` directory mirrors the structure of `R/`. Each test file
corresponds to one source file:

```{r test-structure, eval = FALSE}
tests/testthat/
├── helper.R           # runs before every test file; sets db_path
├── test-utils.R       # tests for axis_vars
├── test-MovieData.R   # tests for MovieData R6 class
├── test-mod_filters.R # tests for mod_filters UI and server
└── test-mod_plot.R    # tests for mod_plot UI and server
```

`helper.R` contains setup code that testthat automatically sources before any
test file runs:

```{r helper, eval = FALSE}
db_path <- system.file("extdata/movies.db", package = "movexplR6")
```

Tests that need the database guard with `skip_if(db_path == "", ...)` so they
are cleanly skipped if the package has not been installed yet (for example, in
a fresh CI environment before `devtools::install()`).

## Development workflow

The standard `devtools` workflow applies without modification:

```{r workflow, eval = FALSE}
# Load all R/ files into the current session (no install needed)
devtools::load_all()

# Re-generate NAMESPACE and man/ from roxygen2 tags
devtools::document()

# Run all tests in tests/testthat/
devtools::test()

# Build and install the package
devtools::install()

# Build all vignettes
devtools::build_vignettes()

# Run the app
launch_app()
```

The key insight is that `devtools::load_all()` + `launch_app()` gives a fast
edit-reload-test loop: edit a file in `R/`, call `load_all()` in the console,
and run the app again — no install required.
